---
title: "Homework 3"
author: Michelle Lui
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
        fig.width = 6,
    fig.asp = .6,
    out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>%
  count(aisle) %>%
  filter(n>10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Apples vs ice cream..

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

###Problem 2

Import, read, and tidy the dataset.

```{r}
accel_df = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "min_of_day",
    names_prefix = "activity_",
    values_to = "activity_count") %>%
  mutate(
    day_type = 
      case_when(
        day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
        day %in% c("Saturday", "Sunday") ~ "weekend")) %>%
  mutate(
    day = as.factor(day),
    day_type = as.factor(day_type),
    min_of_day = as.numeric(min_of_day)) %>%
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  arrange(day)
```
In this dataset,  there are `r ncol(accel_df)` variables, consisting of `r names(accel_df)`. There are `r nrow(accel_df)` observations in this dataset. 

Aggregate across minutes to create a total activity for each day. Create table that shows this. 

```{r}
accel_df %>%
  group_by(week,day) %>%
  summarize(activity_count = sum(activity_count)) %>%
  knitr::kable()
```
For the last few weeks in the datset, it appears that there is very low activity on Saturdays compared to the rest of the days. Activity also seems to increase generally from Mondays through Fridays. The days with the highest values tend to be Fridays and Sundays. 


Make plot that shows 24-hour activity time courses for each day and use color to indicate day of the week. 

```{r}
accel_df %>%
  select(day_id, day, min_of_day, activity_count) %>%
  group_by(day,min_of_day) %>%
  summarize(day_id = day_id, mean_activity_count = mean(activity_count, na.rm = TRUE)) %>%
  ggplot(aes(x = min_of_day, y = mean_activity_count, group = day_id, color = day)) +
    geom_point() +
    geom_path()+
    labs(title = "Average activity count per minute for 35 days", x = "Minute of the day", y = "Average activity count") 
```
For this graph, I took the average of each activity county for each minute of each weekday (Monday, Tuesday, etc.), and plotted that against minutes of the day, and connected the lines based off of each of the 35 days in order to make it easier to interpret. 

Based off of this graph I can see that Sundays and Fridays have the highest activity of the week, with Sundays having higher activity in the middle of the day, and Fridays having higher activity in the evening. There is low activity in the first third of the day, with activity also decreasing at the end of the day. Mondays through Thursdays and Saturdays all seem to have relatively lower activity compared to Fridays and Saturdays. 

###Problem 3

Import and clean data. Convert precipitation, snow depth, and snowfall to inches, and tmax and tmin to degrees C.
```{r}
library(p8105.datasets)
data("ny_noaa")
summary(ny_noaa)
```
In this dataset there are `r nrow(ny_noaa)` observations, and `r ncol(ny_noaa)` variables, which consist of `r names(ny_noaa)`. There is quite of bit of missing data, with prcp missing 145838 observations, snow missing 381221, and snwd missing 591786. Id, tmax, tmin are in the character format, date is in the date format, and prcp, snow, and snwd are in the integer format. 


```{r}
ny_noaa_df = 
  ny_noaa %>%
    separate(date, sep="-", into = c("year", "month", "day")) %>%
    mutate(
      tmax = as.numeric(tmax),
      tmin = as.numeric(tmin),
      tmax = tmax / 10,
      tmin = tmin / 10,
      prcp = prcp * .003937,
      snow = snow * .03937,
      snwd = snwd * .03937) 
 
sort_df = 
  ny_noaa_df %>%
  count(snow, sort = TRUE)
```
For this dataset, I divided the tmin and tmax by 10, so that temperature would be reported as degrees Celsius rather than tenths of a degree Celsius. I also multiplied snow and snwd by .03937, because this is the conversion factor from mm to inches, and I believe rain and snow are usually reported in inches. I multiplied prcp by .003937 because it was originally reported as tenths of a mm, and I wanted to report it as inches. 

For snowfall, the most commonly observed values are NA and 0.00; this is because for most days of the year it is not snowing in New York State, so the amount of snowfall in inches would be 0, or it wouldn't be given at all (NA).


Make a two-panel plot showing the average max temperature in January and in July for each station across the years.

```{r}
plot1_df = 
ny_noaa_df %>%
  group_by(id, year, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>%
  subset(month %in% c("01", "07")) %>%
  drop_na(mean_tmax) %>%
  mutate(
    year = as.numeric(year)
  )

ggplot(plot1_df, aes(x = year, y = mean_tmax, group = id, color = id)) +
  geom_point() +
  geom_path() +
  facet_grid(~ month) + 
  labs(title = "Average maximum temperature for Janurary and July per year", x = "year", y = "Average max temperature (C)") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_continuous(breaks = plot1_df$year)
```
In this 2 panel plot, it is evident that the max temperatures in July (range ~ 20 to 33 degrees C) are much higher than that of January (range ~ -10 to 10 degrees C). There are no particular trends visible within the plots for each month besides that there seems to be peaks and dips between the years quite frequently. Some outliers that are visible January are: one station in 1982 had a very low temperature ~ -14 degrees C, and one in 2005 had a temperature of ~ -12 degrees C. For July, some obvious outliers were one station in 1988 had a lower temperature of ~14 degrees C, and the different stations in the years 1984, 2004, and 2007 all experienced a slightly lower temperature of around ~19 degrees C. 


Make a two panel plot showing tmax vs tmin for the full dataset.

```{r}
plot2_df = 
  ny_noaa_df %>%
  drop_na(tmax, tmin)

plot2 = 
ggplot(plot2_df, aes(x = tmin, y = tmax)) +
      labs(title = "Minimum temperature vs maximum temperature", x = "minimum temperature (C)", y = "maximum temperature (C)") +
         geom_hex()
```

Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
plot3_df = 
  ny_noaa_df %>%
  subset(0 < snow & snow < 100)

plot3 = 
ggplot(plot3_df, aes(x = year, y = snow)) +
  geom_violin() +
  labs(title = "Distribution of snowfall(in) per year", x = "year", y = "snowfall (in)") +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```

Patch them together!

```{r}
library(patchwork)
plot2 / plot3
```

