---
title: "Homework 3"
author: Michelle Lui
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
        fig.width = 6,
    fig.asp = .6,
    out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>%
  count(aisle) %>%
  filter(n>10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Apples vs ice cream..

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

###Problem 2

Import, read, and tidy the dataset.

```{r}
accel_df = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "min_of_day",
    names_prefix = "activity_",
    values_to = "activity_count") %>%
  mutate(
    day_type = 
      case_when(
        day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
        day %in% c("Saturday", "Sunday") ~ "weekend")) %>%
  mutate(
    day = as.factor(day),
    day_type = as.factor(day_type),
    min_of_day = as.numeric(min_of_day)) %>%
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))

```


Aggregate across minutes to create a total activity for each day. Create table that shows this. 

```{r}
table1 = 
accel_df %>%
  group_by(week,day) %>%
  summarize(activity_count = sum(activity_count)) %>%
  knitr::kable()
```

Make plot that shows 24-hour activity time courses for each day and use color to indicate day of the week. 

```{r}
ggplot(accel_df, aes(x = min_of_day, y = activity_count, color = day)) +
  stat_smooth()+
  geom_point() +
  geom_line()
```



